\documentclass[12pt,a4paper]{article}
\usepackage[margin=2.5cm]{geometry}
\usepackage{graphicx} % Required for inserting images
\usepackage{float}
\usepackage{subcaption}
\usepackage{booktabs}
\usepackage{tikz}
\usetikzlibrary{positioning}


\title{4B23: Optical Network Design}
\author{Joe Leacy}

\begin{document}

\maketitle

\section{Introduction}

\begin{figure}[H]
\centering
\begin{subfigure}[b]{0.45\textwidth}
\centering
\begin{tikzpicture}[scale=1.2]
    % Define node positions (clockwise from bottom)
    \node[circle, draw, fill=blue!20, minimum size=0.8cm, label=below:London] (London) at (0, -2) {};
    \node[circle, draw, fill=blue!20, minimum size=0.8cm, label=left:Birmingham] (Birmingham) at (-1.5, -0.5) {};
    \node[circle, draw, fill=blue!20, minimum size=0.8cm, label=left:Manchester] (Manchester) at (-1.5, 1.5) {};
    \node[circle, draw, fill=blue!20, minimum size=0.8cm, label=above:Glasgow] (Glasgow) at (0, 3) {};
    \node[circle, draw, fill=blue!20, minimum size=0.8cm, label=right:Leeds] (Leeds) at (1.5, 0.5) {};
    
    % Draw edges
    \draw[thick] (London) -- (Birmingham);
    \draw[thick] (London) -- (Leeds);
    \draw[thick] (Birmingham) -- (Leeds);
    \draw[thick] (Birmingham) -- (Manchester);
    \draw[thick] (Manchester) -- (Leeds);
    \draw[thick] (Manchester) -- (Glasgow);
    \draw[thick] (Glasgow) -- (Leeds);
\end{tikzpicture}
\caption{Network 1}
\end{subfigure}
\hfill
\begin{subfigure}[b]{0.45\textwidth}
\centering
\begin{tikzpicture}[scale=1.2]
    % Define node positions (clockwise from bottom)
    \node[circle, draw, fill=blue!20, minimum size=0.8cm, label=below:London] (London) at (0, -2) {};
    \node[circle, draw, fill=blue!20, minimum size=0.8cm, label=left:Birmingham] (Birmingham) at (-1.5, -0.5) {};
    \node[circle, draw, fill=blue!20, minimum size=0.8cm, label=left:Manchester] (Manchester) at (-1.5, 1.5) {};
    \node[circle, draw, fill=blue!20, minimum size=0.8cm, label=above:Glasgow] (Glasgow) at (0, 3) {};
    \node[circle, draw, fill=blue!20, minimum size=0.8cm, label=right:Leeds] (Leeds) at (1.5, 0.5) {};
    
    % Draw edges
    \draw[thick] (London) -- (Birmingham);
    \draw[thick] (London) -- (Leeds);
    \draw[thick] (London) -- (Manchester);
    \draw[thick] (Birmingham) -- (Manchester);
    \draw[thick] (Manchester) -- (Glasgow);
    \draw[thick] (Glasgow) -- (Leeds);
\end{tikzpicture}
\caption{Network 2}
\end{subfigure}
\caption{Comparison of two network topologies with 5 UK cities.}
\label{fig:networks}
\end{figure}

\section{Methodology}
\subsection{Routing}
The first step was to determine the proportion of capacity that should be allocated to each route joining two nodes in the network. The number of channels available to each link is approximately 
\begin{equation}
N_i = N_{busiest} \times \frac{C_i}{C_{busiest}}
\label{eq:num_channels}
\end{equation}
where \(N\) is the number of channels and \(C\) is the capacity of a link. Hence an optimal routing will maximize the total channels (and thus total capacity) by minimizing the usage of the busiest link. This can be framed as the multi-commodity flows problem and solved with linear programming given the demands between nodes and the maximum capacities of each link. The result is the relative usage of every route through the network.
\subsection{Channel Assignment}
Equation \ref{eq:num_channels} indicates that the maximum number of channels should be assigned to the busiest link. Here the EDFA limits this to 25. Channels can then be assigned to each route passing through the busiest link in proportion to their usages. One of these routes is then used as a reference to assign channels to those not passing through the busiest link as
\begin{equation}
N_i = N_{ref} \times \frac{C_i}{C_{ref}}.
\end{equation}
A final check is then performed to ensure every link has at most 25 channels before computing the total number of channels as the sum of the channels assigned to each route.
\subsection{NSR Calculation}
The NSR in decibels over an integer number of spans is given by
\begin{equation}
    NSR_{dB} = NF + 10 \log_{10}(N_s) + \alpha L_s + 10 \log_{10}(R_sh\nu \times 10^3) - P_{TX,dBm} + 10 \log_{10}(1.5)
    \label{eq:nsr_spans}
\end{equation}
where \(NF\) is the noise figure of the EDFA, \(N_s\) is the number of spans, \(\alpha\) is the fiber attenuation in dB/km, \(L_s\) is the span length in km, \(R_s\) is the symbol rate, and \(P_{TX, dBm}\) is the transmitted power in dBm. The final term is due to non-linear noise being half that of amplifier noise at the optimum launch power. This assumes a brick-wall shaped pulse and that the amplifer gain in dB is \(\alpha L_s\). In practice the link length will not be an integer number of spans so the additional NSR of the remaining length must be included. The NSR due to non-linear effects in the excess fiber is given by
\begin{equation}
    NSR_{add} = \frac{N_{NLI}}{G_{TX}} = C_{NLI} G_{TX}^2
\end{equation}
where \(C_{NLI}\) is the non-linear coefficient of the excess fiber and \(G_{TX}\) is the PSD of the signal at the start of the excess fiber. An additional amplifier may be needed at the end of the link, in which case the NSR is further increased by
\begin{equation}
    NSR_{amp} = \frac{N_{ASE}}{(10^\frac{-\alpha L_x}{10}g) \times G_{TX}}
\end{equation}
where \(N_{ASE}\) is the PSD of the amplifier noise, \(L_x\) is the excess length in km and \(g\) is the linear amplifier gain. These NSRs can be added in the linear domain, noting that the NSR of the transceiver itself is also included.

\subsection{Determining Amplifier Gain and Span Length}
The amplifiers used had an unsaturated gain \(G_0\) of 25 dB and a total saturation output power across all channels \(P_{sat, dBm}^{out}\) of 23 dBm. The true gain of the amplifier in dB is
\begin{equation}
    G = G_0 - 10 \log_{10} (1 + \frac{P_{out}}{P_{sat}})
    \label{eq:gain}
\end{equation}
Where \(P_{out}\) is the total output power across all channels in mW, and \(P_{sat} = \frac{1}{\ln 2} P_{sat}^{out}\) in mW. Equation \ref{eq:nsr_spans} is valid for the optimum launch power over one channel \(P_{TX, opt} = R_s\sqrt[3]{\frac{N_{ASE}}{C_{NLI}}}\) which can be computed iteratively by first assuming \(G = G_0\), finding the total output power, then updating \(G\) using Equation \ref{eq:gain} and repeating until \(G\) and \(P_{TX, opt}\) converge. Then span length \(L_s = \frac{G}{\alpha}\). Equation \ref{eq:gain} is also used to find the gain of amplifiers added at the end of a link, although there is no iteration here as the excess length and transmitted power are already determined.
\subsection{Validating a Network}
A link in the network will be valid if the received power is greater than the minimum required for the transceiver type being used and the NSR is lower than the maximum the transceiver will permit. First the received power is checked without an additional amplifier, if it is too low the amplifier is added and the total NSR is recalculated. Then the NSR is checked, if it is below the maximum NSR and the received power is high enough then the combination of fiber and transceiver type is valid for the link.
To avoid the practical issues associated with changing fiber type between spans, only one fiber type may be used throughout the network. Transceivers are able to operate in four modes with different data rates but the channels were assigned assuming the same transceiver mode throughout for simplicity, hence only one mode will be used in the network. Using these constraints, a combination of fiber and transceiver mode is valid for the network if it is valid for all of its links.
\subsection{Optimization and Measuring Resilience}
The optimum valid combination of fiber and transceiver for a network was chosen to be that which maximized the total capacity = total channels \(\times\) data rate per channel. If multiple combinations gave the maximum capacity then the one with the fewest amplifiers was chosen. The resilience of each network was determined by its lowest total capacity when any one of its links were broken as well as the corresponding change in capacity available for each node.

\section{Results}


\end{document}